---
title: "DS8004 - Assignment 2"
author: "Najlis, Bernardo"
date: "February 21, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(HMM)
library(gtools)
```

## Hidden Markov Models - Forward Algorithm Implementation

You will submit a piece of code (in language of your choice java/python/c#/R) that implements Forward Algorithm for Hidden Markov Models.

Hint: The best description of the algorithm I could find is in Section III of this paper.


Lawrence R. Rabiner, A Tutorial on Hidden Markov Models and Selected Applications in Speech Recognition. Proceedings of the IEEE, 77 (2), p. 257–286, February 1989. 10.1109/5.18626

### Deliverable:

A. A script that requires three inputs: 

  1) State Transition Matrix (with state names) 
  
  2) Emission Matrix (with symbol names) 
  
  3) Initial State Distribution

B. It requires a sequence of observations vector

C. It output probability of the observations vector given the HMM (specified in A)

  1) Using Exhaustive search

  2) Using Forward Algorithm 

## Resolution

### Forward Algorithm

First we will create the state transition matrix, emission matrix and initial state distribution variables.

```{r cars}
states <- c("rain", "sun")
states_probabilities <- c(0.4, 0.2, 0.6, 0.8)
symbols <- c("happy", "grumpy")
emission_probabilities <- c(0.4, 0.9, 0.6, 0.1)
initial_probabilities <- c(0.5, 0.5)
observations = c("happy", "grumpy", "happy")

state_transition_matrix <- matrix(states_probabilities, nrow=length(states), ncol=length(states), dimnames=list(states, states))
emission_matrix <- matrix(emission_probabilities, nrow=length(states), ncol=length(symbols), dimnames=list(states, symbols))
initial_state_distribution <- matrix(initial_probabilities, nrow=1, ncol=length(states), dimnames=list(c("intial probabilities"),states))
```

Let's look at the values on these input matrices:
 
```{r}
state_transition_matrix
emission_matrix
initial_state_distribution

fwd_algo_alpha = array(0, c(length(states), length(observations)), dimnames = list(states, index = 1:length(observations)))

for (state_num in 1:length(states)) {
  fwd_algo_alpha[state_num,1] <- initial_state_distribution[state_num] * emission_matrix[state_num,observations[1]]
}

# every next time
for (obs_num in 2:length(observations)) { # for each observation
  for (state_num in 1:length(states)) { # for each state
    temp <- 0
    for (prev_state_num in 1:length(states)){ # prior state for each current state
      temp <- temp + state_transition_matrix[prev_state_num,state_num] * fwd_algo_alpha[prev_state_num,obs_num-1] # sigma(alpha * transition matrix)
    }
    fwd_algo_alpha[state_num,obs_num] <- temp * emission_matrix[state_num,observations[obs_num]] # updated probabilities for each state
  }
}

```

Building an HMM model using the "HMM" package to compare...

```{r, warning=FALSE}

test_hmm <- initHMM(States = states, 
                    Symbols = symbols, 
                    startProbs = initial_probabilities, 
                    transProbs = state_transition_matrix, 
                    emissionProbs = emission_matrix)
test_hmm


logfwdprobs <- forward(test_hmm, observations)
```
These are the forward probabilities based on the HMM package:
```{r}
print(exp(logfwdprobs))
```
These are the forward probabilities based on the implemented forward algorithm:
```{r}
print(fwd_algo_alpha)
```
